---
title: "Portfolio 2"
author: "Veethacheng Thao"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

To build upon the first part of the portfolio we will load it.
```{r echo=FALSE, message=FALSE, error=FALSE, warning=FALSE, results='hide'}
install.packages("tidyverse", repos = "http://cran.us.r-project.org")
library("tidyverse")
library("knitr")
purl("portfolio.Rmd", output = "part1.r")
source("part1.r")
```

The new source of data will now be json which can be found at https://github.com/dariusk/corpora/blob/master/data/medicine/drugs.json.
To load the data we use rjson package.
```{r}
library("rjson")
json_file <- "https://raw.githubusercontent.com/dariusk/corpora/master/data/medicine/drugs.json"
json_data <- fromJSON(file=json_file)
```

Putting the list of pharmaceutical drugs to a dataframe.
```{r}
pharmaceutical <- as.data.frame(json_data["drugs"])
```

Now since we have our two datasets I want to know if any of the drugs that were being abuse from my previous dataset appear as any of these pharmaceutical drugs.
But first we need to fix up the dataframes so they can be easily compared.
```{r}
library("dplyr")
colnames(drugs)
only_drugs <- drugs %>% select(-Age, -Gender, -Education, -Country, -Ethnicity)
abused_drugs <- as.data.frame(colnames(only_drugs))
colnames(abused_drugs)[colnames(abused_drugs)=="colnames(only_drugs)"] <- "abused_drug"
```
Now to compare we use the %in%.
```{r}
abused_drugs %in% pharmaceutical
```
The results is false just as i hope it to be since all the abused drugs were mostly hard drugs.

Next it would be interesting to see if age has any correlation into using some of these hard drugs.
```{r}
ll <- lm(data = drugs, Age ~ Coke + Meth + Crack + Mushrooms)
summary(ll)
```